---
title: "ASSIGNMENT 2"
author: "LEVIN S"
date: "2022-11-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
install.packages("nycflights13")
install.packages("tidyverse")
library(nycflights13)
library(tidyverse)
```





1.	Sort flights to find the fastest flights

```{r}
arrange(flights,arr_time)

flights %>% mutate(travel_time = ifelse((arr_time - dep_time < 0), 
                                        2400+(arr_time - dep_time),
                                        arr_time - dep_time)) %>% 
  arrange(travel_time) %>% select(arr_time, dep_time, travel_time)

arrange(flights, (arr_time - dep_time))
```
 
2.	Compare dep_time, sched_dep_time, and dep_delay. How would you expect those three numbers to be related?

Answer:
 I would expect that is "dep_delay" is equal to the difference between "sched_dep_time" and "dep_time"


```{r}
flights %>% 
  mutate (dep_time = (dep_time %/% 100) * 60 + (dep_time %% 100),
         sched_dep_time = (sched_dep_time %/% 100) * 60 + (sched_dep_time %% 100),
         arr_time = (arr_time %/% 100) * 60 + (arr_time %% 100),
         sched_arr_time = (sched_arr_time %/% 100) * 60 + (sched_arr_time %% 100)) %>%
  transmute (near((sched_dep_time + dep_delay) %% (60*24), dep_time, tol=1))

select(flights,dep_time,sched_dep_time,dep_delay)

```
3.	Brainstorm at least 5 different ways to assess the typical delay characteristics of a group of flights. Consider the following scenarios:
a.	A flight is 15 minutes early 50% of the time, and 15 minutes late 50% of the time.
b.	A flight is always 10 minutes late.
c.	A flight is 30 minutes early 50% of the time, and 30 minutes late 50% of the time.
d.	99% of the time a flight is on time. 1% of the time it’s 2 hours late.

Answer:

```{r}
not_cancelled <- flights %>%
  filter(!is.na(air_time))
###Arrival Time is More Important so sort that first
not_cancelled %>%
  group_by(tailnum) %>%
  mutate(
    count = n(),
    median_arr_delay = median(arr_delay),
    median_dep_delay = median(dep_delay)
  ) %>%
  filter(count > 30) %>%
  arrange(median_arr_delay, median_dep_delay)


###More than 10 mins Late
not_cancelled %>%
  group_by(tailnum, origin, dest) %>%
  summarise(
    count = n(),
    arr_delay_10_c = sum(arr_delay > 10),  
    arr_delay_10_p = mean(arr_delay > 10),  
    dep_delay_10_c = sum(dep_delay > 10),
    dep_delay_10_p = mean(dep_delay > 10)  
    ) %>%
  filter(count > 20) %>%
  arrange(desc(arr_delay_10_p))


### Best flights under 30 mins early or late criteria sorted 
not_cancelled %>%
  group_by(tailnum) %>%
  mutate(
    count = n(),
    arr_30_early = mean(arr_delay < -30),
    dep_30_early = mean(dep_delay < -30),
    arr_30_late  = mean(arr_delay > 30),
    dep_30_late = mean(dep_delay > 30)
  ) %>%
  filter(count > 20) %>%
  arrange(desc(arr_30_early), desc(dep_30_early), arr_30_late, dep_30_late) %>%
  select(dest)

```
According to the scenarios and coding ,departure delay is important


4.	Which plane (tailnum) has the worst on-time record?

Answer:

```{r}
flights %>%
  group_by(tailnum) %>%
  summarise(prop_on_time = sum(arr_delay <= 30 & !is.na(arr_delay))/n(),
            mean_arr_delay = mean(arr_delay, na.rm=TRUE),
            flights = n()) %>%
  arrange(prop_on_time, desc(mean_arr_delay))
flights %>%
  group_by(tailnum) %>%
  filter(all(is.na(arr_delay))) %>%
    tally(sort=TRUE)

arrange(flights,desc(arr_delay),desc(dep_delay))
```
5.	Brainstorm as many ways as possible to select dep_time, dep_delay, arr_time, and arr_delay from flights.


Answer:

```{r}
select(flights, dep_time,  dep_delay, arr_time, arr_delay)
select(flights,4,6,7,9)
select(flights,starts_with("dep_"),starts_with("arr_"))
select(flights, "dep_time", "dep_delay", "arr_time", "arr_delay")

flights %>% select(ends_with("time"),  ends_with("delay")) %>% select(-c(starts_with("dep"), starts_with("arr")))


#flights %>% select(dep_time,  dep_delay, arr_time, arr_delay)
#flights %>% select_("dep_time",  "dep_delay", "arr_time", "arr_delay")
#flights %>% select_(.dots=c("dep_time",  "dep_delay", "arr_time", "arr_delay"))
# fancier ways
#flights %>% select(dep_time:arr_delay, -c(contains("sched")))


#flights %>% select(contains("dep"), contains("arr"), -contains("sched"), -carrier)
#flights %>% select(matches("^dep|arr_delay|time$"))
#flights %>% select(matches("^dep|^arr"))
#flights %>% select(matches("^dep|^arr.*time$|delay$"))
#flights %>% select(matches("^dep|^arr_time$|delay$"))
```





